---
title: "Data for: "
author: "Nicholas S. Marzolf"
date: "Updated March 13, 2023"
output: 
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

RMarkdown file to accompany *On the breakdown of woody debris across a groundwater gradient in Neotropical streams, Costa Rica,* submitted to the journal *Freshwater Science.*

To recreate this HTML, load the data files into a folder named 'Data' at the same path as the .rmd file, open the .rmd file, and click 'Knit'.

# Load packages

```{r load-packages, echo = TRUE, message=FALSE, warning = FALSE}
# data manipulation
library(tidyverse)
library(readxl)
library(dplyr)
library(forcats)
library(purrr)

# plotting
library(ggplot2)
library(ggeffects)
library(grid)
library(ggpubr)
library(ggrepel)
library(ggcorrplot)
library(lemon)
source('C:/Users/Nick Marzolf/Desktop/Research/R code/theme_nick.R')
theme_set(theme_nick())

# statistics
library(car)
library(drc)
library(nlme)
library(nlstools)
library(qpcR)

# spatial
library(sf)
library(raster)
library(spData)
library(cowplot)
library(leaflet)
library(spDataLarge)

# community
library(vegan)
library(pander)
library(lattice)
library(permute)

```

# Load data

```{r load-data, warning=FALSE, echo=TRUE}
chem <- readxl::read_excel('Data/LT_CWD_datasheet.xlsx',
                           sheet = 'Chemistry') 

cwd <- readxl::read_excel('Data/LT_CWD_datasheet.xlsx',
                          sheet = 'Sheet1')

table_s1 <- readxl::read_excel('Data/LT_CWD_datasheet.xlsx',
                               sheet = 'Wood Morphology')
readr::write_csv(table_s1,
                 'Data/table_s1.csv')
```

# Stream chemistry for 5 sites

```{r chemistry, warning=FALSE, echo=TRUE}
chem <- chem %>% 
  dplyr::rename(srp = `SRP (ug/L)`,
                no3_n = `NO3-N (ug/L)`,
                nh4_n = `NH4-N (ug/L)`) %>% 
  dplyr::mutate(site = case_when(Site == 'Arb' ~ 'Arboleda-30',
                                 Site == 'Sur30' ~ 'Sura-30',
                                 Site == 'Tito60' ~ 'Saltito-60',
                                 Site == 'Piper' ~ 'Piper-30',
                                 Site == 'Tac' ~ 'Taconazo-30'))

# calculate mean of all measurements
chem_sum <- chem %>% 
  dplyr::group_by(site) %>%
  dplyr::summarise(dplyr::across(.cols = 2:7,
                                 .fns = mean, 
                                 na.rm = TRUE))

# calculate SD from stream chemistry data
chem_sd <- chem %>% 
  dplyr::group_by(site) %>%
  dplyr::summarise(dplyr::across(.cols = 2:7,
                                 .fns = sd, 
                                 na.rm = TRUE))

# create object that sorts sites by decreasing mean conductivity
sites <- chem_sum %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean_cond = mean(Cond, na.rm = TRUE)) %>%
  dplyr::arrange(desc(mean_cond)) 

# create a characeter vector of sites in order
sites <- as.character(sites$site)


# calculate DIN and N:P ratio
chem_sum <- chem_sum %>%
  dplyr::mutate(din = no3_n + nh4_n,
                n_p = (din/14.0067)/(srp/30.973762))


# re-level the site factor
chem_sum$site <- forcats::fct_relevel(chem_sum$site,
                                      sites)

table1 <- dplyr::arrange(chem_sum, 
                         desc(Cond))
table1
```

# Figure 1: Map

```{r map, warning=FALSE, echo=TRUE}
streams_gps <- readr::read_csv('C:/Users/Nick Marzolf/Desktop/NCSU/STREAMS/La Selva GIS data/LTREB Data/GPS sites.csv')

cwd_coords <- streams_gps %>% 
  dplyr::filter(Site %in% sites_long) %>% 
  dplyr::mutate(hyphen_site = gsub(' ', '-', Site)) %>% 
  dplyr::mutate(Site = ifelse(Site == 'Piper',
                              'Piper-30',
                              hyphen_site),
                Elevation = ifelse(Site == 'Piper-30',
                                   30,
                                   Elevation)) %>% 
  dplyr::select(-hyphen_site) %>% 
  sf::st_as_sf(., 
               coords = c('Long', 'Lat'),
               crs = st_crs(4326))

cwd_coords$Site <- forcats::fct_relevel(cwd_coords$Site,
                                        c('Arboleda-30', 'Sura-30', 'Saltito-60',
                                          'Piper-30', 'Taconazo-30'))

# La Selva boundary
lsbs <- sf::st_read(dsn = 'Data/Spatial/laselvaboundary.shp')

# Streams at La Selva shapefile
streams <- sf::st_read(dsn = 'Data/Spatial/streamsclip.shp')

# shapefile for Costa Rica
cr <- world %>%
  dplyr::filter(name_long == 'Costa Rica')

# make site map, with 5 streams
sites_map <- ggplot()+
  geom_sf(data = lsbs, 
          fill = 'white',
          color = 'black')+
  geom_sf(data = streams,
          color = 'blue')+
  geom_sf(data = cwd_coords,
          size = 5,
          aes(color = Site))+
  coord_sf(xlim = c(-84.05, -83.98),
           ylim = c(10.395, 10.45),
           crs = 4326)+
  scale_color_viridis_d()+
  ggspatial::annotation_scale(location = "bl",
                              bar_cols = c("black", "white"))+
  ggspatial::annotation_north_arrow(location = "tl", 
                                    which_north = "true",
                                    style = ggspatial::north_arrow_minimal(line_col = "black"))

# create a centroid for La Selva
centroid <- sf::st_centroid(lsbs[1,])

# create the inset map
inset <- ggplot()+
  geom_sf(data = cr)+
  geom_sf(data = centroid)+
  scale_x_continuous(breaks = -84)+
  scale_y_continuous(breaks = 10)

# make the full figure
fig1 <- ggdraw()+
  draw_plot(sites_map)+
  draw_plot(inset, x = .6, y = 0.1, height = 0.25, width = 0.25)+
  theme()

# save the figure
ggsave(plot = fig1,
       filename = 'Figures/fig1_ggplot.png',
       dpi = 600, 
       width = 9.85, height = 7)

```

# Calculate wood decomposition rates

```{r kCWD, warning=FALSE, echo=TRUE}
# begin cleaning data
cwd <- cwd %>%
  dplyr::mutate(site = case_when(Site == 'Arb' ~ 'Arboleda-30',
                                 Site == 'Sur30' ~ 'Sura-30',
                                 Site == 'Tito60' ~ 'Saltito-60',
                                 Site == 'Piper' ~ 'Piper-30',
                                 Site == 'Tac' ~ 'Taconazo-30')) %>% 
  dplyr::filter(Flag == 0) %>%
  dplyr::select(site, 
                month = `Collection Month`,
                rep = Rep,
                init_mass = `initial CWD mass (g)`,
                dry_mass = `CWD Pack Dry Mass (g)`,
                init_den = `init wood density (g/cm3)`,
                fin_den = `final wood density (g/cm3)`) %>%
  dplyr::filter(month < 24)

# re-level site factor based on conductivity
cwd$site <- forcats::fct_relevel(cwd$site, 
                                 sites)

# calculate % dry mass remaining
cwd_calc <- cwd %>%
  dplyr::filter(site != 'Sac') %>%
  dplyr::mutate(percent_mass = (dry_mass/init_mass)*100)

# dry mass of each pack to be used in the bug analysis
final_dry_mass <- cwd_calc %>% 
  dplyr::select(site, month, rep, dry_mass) %>% 
  dplyr::filter(month != 0)

# calculate decay rates using ANCOVA
k_cwd_int <- lm(data = cwd_calc,
                log(percent_mass) ~ month * site)

summary(k_cwd_int)
anova(k_cwd_int)
car::Anova(k_cwd_int, type = 'III')

table2 <- cwd_calc %>%
  dplyr::filter(site != 'Sac') %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(
    int = round((coef(lm(log(percent_mass) ~ month))[1]), 3),
    k_yr = (round((coef(lm(log(percent_mass) ~ month))[2])*12, 4))*-1,
    error = summary(lm(log(percent_mass) ~ month))$coefficient[3],
    #df = summary(lm(log(percent_mass) ~ month))$fstatistic,
    r2 = round(summary(lm(log(percent_mass) ~ month))$r.squared, 3),
    p = anova(lm(log(percent_mass) ~ month))$'Pr(>F'[1])
table2


cwd_plot <- dplyr::left_join(cwd_calc, 
                             table2, 
                             by = 'site')

fig2 <- ggplot(data = cwd_plot,
               aes(y = log(percent_mass),
                   x = month))+
  geom_point(aes(color = site))+
  geom_smooth(aes(color = site),
              method = 'lm',
              se = FALSE)+
  lims(y = c(3.4, 5))+
  geom_text(data = table2 %>% 
              dplyr::mutate(label = paste("italic(k) == ", k_yr)),
            aes(x = 2.5, y = 3.7,
                label = label),
            parse = TRUE)+
  geom_text(data = table2 %>% 
              dplyr::mutate(label = paste("italic('R'^2) == ", r2)),
            aes(x = 2.5, 
                y = 3.45,
                label = label),
            parse = TRUE,
            inherit.aes = FALSE)+
  facet_grid(. ~ site)+
  labs(x = 'Months in stream',
       y = 'ln(% DM remaining)')+
  scale_color_viridis_d()+
  scale_x_continuous(breaks = c(0, 1, 3, 6, 12))+
  theme_classic()+
  theme(legend.position = 'none',
        panel.grid = element_blank(),
        strip.background = element_blank())
fig2

ggsave(plot = fig2, 
       'Figures/fig2.png',
       dpi = 600, 
       width = 10,height = 3)

cwd_calc %>% 
  group_by(site, month) %>% 
  summarise(mean_per_mass = mean(percent_mass, na.rm = TRUE))

```

# Merge chemistry with decay rates

```{r decay-chemistry, warning=FALSE, echo=TRUE}
merged <- table2 %>%
  dplyr::select(site, k_yr, error) %>% 
  dplyr::right_join(chem_sum, 'site') %>%
  # dplyr::select(-Temp) %>%
  tidyr::pivot_longer(srp:n_p)

names_long <- c(`Cond` = 'Conductivity (µS/cm)',
                `din` = 'DIN (µg/L)',
                `n_p` = 'DIN:SRP',
                `nh4_n` = 'NH4-N (µg/L)',
                `no3_n` = 'NO3-N (µg/L)',
                `pH` = 'pH',
                `srp` = 'SRP (µg/L)')

wide <- table2 %>%
  dplyr::select(site, k_yr, error) %>% 
  dplyr::right_join(chem_sum, 'site') %>%
  dplyr::select(-Temp)

cor <- cor(wide %>% 
             dplyr::select(-site, -error) %>% 
             mutate_if(is.character, as.numeric), 
           method = 'spearman')

cor_sub <- cor[-c(2:nrow(cor)), 2:ncol(cor),drop = FALSE]
rownames(cor_sub)[1] <- 'k_yr'
colnames(cor_sub)[1:7] <- c('SRP', 'NO3-N', 'NH4-N', 'pH', 'Cond', 'DIN', 'DIN:SRP')

pmat <- ggcorrplot::cor_pmat(cor,
                             'spearman', alternative = 'two.sided')
pmat_sub <- pmat[-c(2:nrow(pmat)), 2:ncol(pmat),drop = FALSE]

ggcorrplot(cor_sub, lab = TRUE,
           ggtheme = theme_nick(),show.legend = FALSE,p.mat = pmat_sub)



```

```{r old-modeling stufff}
<!-- ## Run linear and non-linear models of kCWD ~ chemistry -->
  
  <!-- ```{r linear-mods, warning=FALSE, echo=TRUE} -->
  
  <!-- vars <- unique(merged$name) -->
  
  <!-- # linear: Y ~ a + bX -->
  
  <!-- # create empty dataframe to populate in the for loop -->
  
  <!-- linear <- data.frame(mod = character(), -->
                              
                              <!--                      var = character(), -->
                              
                              <!--                      a = numeric(), -->
                              
                              <!--                      a_2.5 = numeric(), -->
                              
                              <!--                      a_97.5 = numeric(), -->
                              
                              <!--                      b = numeric(), -->
                              
                              <!--                      b_2.5 = numeric(), -->
                              
                              <!--                      b_97.5 = numeric(), -->
                              
                              <!--                      AIC = numeric(), -->
                              
                              <!--                      p = numeric(), -->
                              
                              <!--                      rse = numeric(), -->
                              
                              <!--                      df = numeric()) -->
  
  <!-- # for loop across each parameter -->
  
  <!-- for(i in 1:length(vars)){ -->
      
      <!--   # which var is being predicted -->
      
      <!--   var <- vars[i] -->
        
        <!--   # run the lm of k ~ chemistry -->
        
        <!--   lm <- lm(data = merged %>% -->
                          
                          <!--              filter(name == var), -->
                          
                          <!--            k_yr ~ value) -->
          
          <!--   # populate the empty data frame -->
          
          <!--   linear <- linear %>%  -->
            
            <!--     add_row( -->
                                
                                <!--       mod = 'Linear',                    # what kind of model -->
                              
                              <!--       var = var,                         # for which variable -->
                              
                              <!--       a = coef(lm)[1],                   # intercept -->
                              
                              <!--       a_2.5 = confint(lm)[1,1],          # intercept low CI -->
                              
                              <!--       a_97.5 = confint(lm)[1,2],         # intercept high CI -->
                              
                              <!--       b = coef(lm)[2],                   # slope -->
                              
                              <!--       b_2.5 = confint(lm)[2,1],          # slope low CI -->
                              
                              <!--       b_97.5 = confint(lm)[2,2],         # slope high CI -->
                              
                              <!--       AIC = AIC(lm),                     # model AIC -->
                              
                              <!--       p = summary(lm)$coefficients[2,4], # p value -->
                              
                              <!--       rse = summary(lm)$sigma,           # RSE -->
                              
                              <!--       df = summary(lm)$df[2]             # df -->
                              
                              <!--     ) -->
            
            <!-- } -->
  
  <!-- ``` -->
  
  <!-- ```{r Mic-Men-mods, warning=FALSE, echo=TRUE} -->
  
  <!-- # empty data frame -->
  
  <!-- micmen <- data.frame(mod = character(), -->
                              
                              <!--                      fit = character(), -->
                              
                              <!--                      var = character(), -->
                              
                              <!--                      a = numeric(), -->
                              
                              <!--                      a_2.5 = numeric(), -->
                              
                              <!--                      a_97.5 = numeric(), -->
                              
                              <!--                      b = numeric(), -->
                              
                              <!--                      b_2.5 = numeric(), -->
                              
                              <!--                      b_97.5 = numeric(), -->
                              
                              <!--                      AIC = numeric(), -->
                              
                              <!--                      p = numeric()) -->
  
  <!-- # for loop for each variable -->
  
  <!-- for(i in 1:length(vars)) { -->
      
      <!--   var <- vars[i] -->
        
        <!--   mm_mod <- try(nls(data = merged %>%  -->
                                   
                                   <!--                       filter(name == var), -->
                                   
                                   <!--                     k_yr ~ SSmicmen(value, a, b)) -->
                               
                               <!--   ) -->
          
          <!--   # fault tolerance, won't work for every parameter -->
          
          <!--   if(inherits(mm_mod, 'try-error')){ -->
              
              <!--     micmen <- micmen %>%  -->
                
                <!--       add_row(fit = 'error', -->
                                     
                                     <!--               var = var, -->
                                     
                                     <!--               mod = 'Michaelis-Menten') -->
                
                <!--     next -->
                
                <!--   } -->
          
          <!--   # populate the data frame -->
          
          <!--   micmen <- micmen %>%  -->
            
            <!--     add_row( -->
                                
                                <!--       mod = 'Michaelis-Menten', -->
                                
                                <!--       fit = 'success', -->
                                
                                <!--       var = var, -->
                                
                                <!--       a = coef(mm_mod)[1], -->
                                
                                <!--       a_2.5 = confint2(mm_mod, level = 0.95)['a',1], -->
                                
                                <!--       a_97.5 = confint2(mm_mod, level = 0.95)['a',2], -->
                                
                                <!--       b = coef(mm_mod)[2], -->
                                
                                <!--       b_2.5 = confint2(mm_mod, level = 0.95)['b',1], -->
                                
                                <!--       b_97.5 = confint2(mm_mod, level = 0.95)['b',2], -->
                                
                                <!--       AIC = AIC(mm_mod), -->
                                
                                <!--       p = summary(mm_mod)$coefficients[8] -->
                                
                                <!--     ) -->
            
            <!-- } -->
  
  <!-- ``` -->
  
  <!-- ```{r logarithmic, warning=FALSE, echo=TRUE} -->
  
  <!-- ## Logarithmic: Y ~ a + b*log(X) -->
  
  <!-- logarithmic <- data.frame(mod = character(), -->
                                   
                                   <!--                           var = character(), -->
                                   
                                   <!--                           a = numeric(), -->
                                   
                                   <!--                           a_2.5 = numeric(), -->
                                   
                                   <!--                           a_97.5 = numeric(), -->
                                   
                                   <!--                           b = numeric(), -->
                                   
                                   <!--                           b_2.5 = numeric(), -->
                                   
                                   <!--                           b_97.5 = numeric(), -->
                                   
                                   <!--                           AIC = numeric(), -->
                                   
                                   <!--                           R2 = numeric(), -->
                                   
                                   <!--                           p = numeric(), -->
                                   
                                   <!--                           rse = numeric(), -->
                                   
                                   <!--                           df = numeric() -->
                                   
                                   <!-- ) -->
  
  <!-- for(i in 1:length(vars)) { -->
      
      <!--   var <- vars[i] -->
        
        <!--   loga <- lm(data = merged %>%  -->
                            
                            <!--                filter(name == var), -->
                            
                            <!--              k_yr ~ log10(value))  -->
          
          <!--   logarithmic <- logarithmic %>%  -->
            
            <!--     add_row( -->
                                
                                <!--       mod = 'Logarithmic', -->
                                
                                <!--       var = var, -->
                                
                                <!--       a = coef(loga)[1], -->
                                
                                <!--       a_2.5 = confint(loga)[1], -->
                                
                                <!--       a_97.5 = confint(loga)[3], -->
                                
                                <!--       b = coef(loga)[2], -->
                                
                                <!--       b_2.5 = confint(loga)[2], -->
                                
                                <!--       b_97.5 = confint(loga)[4], -->
                                
                                <!--       AIC = AIC(loga), -->
                                
                                <!--       R2 = summary(loga)$r.squared, -->
                                
                                <!--       p = summary(loga)$coefficients[8], -->
                                
                                <!--       rse = summary(loga)$sigma, -->
                                
                                <!--       df = summary(loga)$df[2] -->
                                
                                <!--     ) -->
            
            <!-- } -->
  
  <!-- ``` -->
  
  <!-- ```{r logistic, warning=FALSE, echo=TRUE} -->
  
  <!-- ## Logistic: Y ~ 1/1+exp(X) -->
  
  <!-- logistic <- data.frame(mod = character(), -->
                                
                                <!--                        var = character(), -->
                                
                                <!--                        b = numeric(), -->
                                
                                <!--                        b_2.5 = numeric(), -->
                                
                                <!--                        b_97.5 = numeric(), -->
                                
                                <!--                        d = numeric(), -->
                                
                                <!--                        d_2.5 = numeric(), -->
                                
                                <!--                        d_97.5 = numeric(), -->
                                
                                <!--                        e = numeric(), -->
                                
                                <!--                        e_2.5 = numeric(), -->
                                
                                <!--                        e_97.5 = numeric(), -->
                                
                                <!--                        AIC = numeric(), -->
                                
                                <!--                        rse = numeric(), -->
                                
                                <!--                        df = numeric()) -->
  
  <!-- for(i in 1:length(vars)){ -->
      
      <!--   var <- vars[i] -->
        
        <!--   logi <- drm(k_yr ~ value,  -->
                             
                             <!--               data = merged %>%  -->
                             
                             <!--                 filter(name == var), -->
                             
                             <!--               fct = L.3()) -->
          
          <!--   logistic <- logistic %>%  -->
            
            <!--     add_row( -->
                                
                                <!--       mod = 'Logistic', -->
                                
                                <!--       var = var, -->
                                
                                <!--       b = coef(logi)[1], -->
                                
                                <!--       b_2.5 = confint2(logi)['b:(Intercept)',1], -->
                                
                                <!--       b_97.5 = confint2(logi)['b:(Intercept)',2], -->
                                
                                <!--       d = coef(logi)[2], -->
                                
                                <!--       d_2.5 = confint2(logi)['d:(Intercept)',1], -->
                                
                                <!--       d_97.5 = confint2(logi)['d:(Intercept)',2], -->
                                
                                <!--       e = coef(logi)[3], -->
                                
                                <!--       e_2.5 = confint2(logi)['e:(Intercept)',1], -->
                                
                                <!--       e_97.5 = confint2(logi)['b:(Intercept)',2], -->
                                
                                <!--       AIC = AIC(logi), -->
                                
                                <!--       rse = summary(logi)$rseMat[1], -->
                                
                                <!--       df = summary(logi)$rseMat[2]) -->
            
            <!-- } -->
  
  <!-- ``` -->
  
  <!-- ## Combine and evaluate model fits -->
  
  <!-- ```{r model-comps, warning=FALSE, echo=TRUE} -->
  
  <!-- # combine model outputs -->
  
  <!-- aic_all <- rbind(logistic %>%  -->
                          
                          <!--                    dplyr::select(mod, var, AIC) %>%  -->
                          
                          <!--                    mutate(p = NA), -->
                          
                          <!--                  logarithmic %>%  -->
                          
                          <!--                    dplyr::select(mod, var, AIC, p), -->
                          
                          <!--                  linear %>%  -->
                          
                          <!--                    dplyr::select(mod, var, AIC, p), -->
                          
                          <!--                  micmen %>%  -->
                          
                          <!--                    dplyr::select(mod, var, AIC, p)) -->
  
  <!-- best_fits <- aic_all %>%  -->
  
  <!--   arrange(var, -->
                   
                   <!--           AIC) %>%  -->
  
  <!--   group_by(var) %>%  -->
  
  <!--   dplyr::slice_head(n=1) -->
  
  <!-- aic_wts <- data.frame() -->
  
  <!-- for(i in 1:length(vars)) { -->
      
      <!--   use <- vars[i] -->
        
        <!--   df <- aic_all %>%  -->
          
          <!--     filter(!is.na(AIC), -->
                            
                            <!--            var %in% use) %>%  -->
          
          <!--     mutate(delAIC = akaike.weights(AIC)$deltaAIC, -->
                            
                            <!--            weights = akaike.weights(AIC)$weights) -->
          
          <!--   aic_wts <- rbind(aic_wts, df) -->
            
            <!-- } -->
  
  <!-- table3 <- aic_wts %>%  -->
  
  <!--   dplyr::select(var, mod, weights) %>%  -->
  
  <!--   pivot_wider(names_from = var,  -->
                       
                       <!--               values_from = weights) -->
  
  <!-- table3 -->
  
  <!-- ``` -->
  
  <!-- ```{r pred-k, warning=FALSE, echo=TRUE} -->
  
  <!-- # plot decay rates as a function of chemistry -->
  
  <!-- # add lines for SRP (linear), NH4 (M-M), N:P(M-M), and cond (linear) -->
  
  <!-- merged_preds <- data.frame() -->
  
  <!-- for(i in 1:length(best_fits$var)) { -->
      
      <!--   fit <- best_fits[i,] -->
        
        <!--   type <- pull(fit[1]) -->
          
          <!--   var <- pull(fit[2]) -->
            
            <!--   if(type == 'Logistic'){ -->
                
                <!--     logis <- logistic %>%  -->
                  
                  <!--       filter(var == !!var) -->
                  
                  <!--     df <- merged %>%  -->
                    
                    <!--       filter(name == var) %>%  -->
                    
                    <!--       mutate(pred_k = logis$d/(1+exp(logis$b*(value - logis$e))), -->
                                        
                                        <!--              best_fit = type) -->
                    
                    <!--   } # end logistic if statement -->
          
          <!--   if(type == 'Linear') { -->
              
              <!--     line <- linear %>%  -->
                
                <!--       filter(var == !!var) -->
                
                <!--     if(line$p <= 0.05){ -->
                    
                    <!--       df <- merged %>%  -->
                      
                      <!--         filter(name == var) %>%  -->
                      
                      <!--         mutate(pred_k = line$a + (value*line$b), -->
                                            
                                            <!--                best_fit = type) -->
                      
                      <!--     } else { -->
                          
                          <!--       df <- merged %>%  -->
                            
                            <!--         filter(name == var) %>%  -->
                            
                            <!--         mutate(pred_k = NA, -->
                                                  
                                                  <!--                best_fit = NA) -->
                            
                            <!--     } -->
                
                <!--   } # end linear if statement -->
          
          <!--   if(type == 'Michaelis-Menten'){ -->
              
              <!--     mm <- micmen %>%  -->
                
                <!--       filter(var == !!var) -->
                
                <!--     if(mm$p <= 0.05){ -->
                    
                    <!--       df <- merged %>%  -->
                      
                      <!--         filter(name == var) %>%  -->
                      
                      <!--         mutate(pred_k = (mm$a*value)/(value + mm$b), -->
                                            
                                            <!--                best_fit = type) -->
                      
                      <!--     } else { -->
                          
                          <!--       df <- merged %>%  -->
                            
                            <!--         filter(name == var) %>%  -->
                            
                            <!--         mutate(pred_k = NA, -->
                                                  
                                                  <!--                best_fit = NA) -->
                            
                            <!--     } -->
                
                <!--   } # end M-M for loop -->
          
          <!--   if(type == 'Logarithmic'){ -->
              
              <!--     logar <- logarithmic %>%  -->
                
                <!--       filter(var == !!var) -->
                
                <!--     if(logar$p <= 0.05) { -->
                    
                    <!--       df <- merged %>%  -->
                      
                      <!--         filter(name == var) %>%  -->
                      
                      <!--         mutate(pred_k = logar$a + (logar$b*value), -->
                                            
                                            <!--                best_fit = type) -->
                      
                      <!--     } else { -->
                          
                          <!--       df <- merged %>%  -->
                            
                            <!--         filter(name == var) %>%  -->
                            
                            <!--         mutate(pred_k = NA, -->
                                                  
                                                  <!--                best_fit = NA) -->
                            
                            <!--     } -->
                
                <!--   } # end logarithmic if statement -->
          
          <!--   merged_preds <- rbind(merged_preds, df) -->
            
            <!-- } # end for loop -->

<!-- fig3 <- ggplot(merged_preds)+ -->
  
  <!--   geom_point(aes(x = value, -->
                          
                          <!--                  y = k_yr, -->
                          
                          <!--                  color = site), -->
                      
                      <!--              size = 4)+ -->
  
  <!--   geom_line(aes(x = value,  -->
                         
                         <!--                 y = pred_k), -->
                     
                     <!--             size = 1, -->
                     
                     <!--             linetype = 'dashed')+ -->
  
  <!--   ylab(expression(paste(k[wood], ' (', y^-1,')')))+ -->
  
  <!--   facet_wrap(. ~ name, -->
                      
                      <!--              nrow = 2, -->
                      
                      <!--              scales = 'free', -->
                      
                      <!--              labeller = as_labeller(names_long), -->
                      
                      <!--              drop = FALSE)+ -->
  
  <!--   scale_color_viridis_d(name = 'Site', -->
                                 
                                 <!--                         labels = as_labeller(sites_long))+ -->
  
  <!--   theme(axis.title.x = element_blank(), -->
                 
                 <!--         legend.background = element_blank(), -->
                 
                 <!--         legend.box.background = element_rect(colour = "black")) -->
  
  <!-- shift_legend2 <- function(p) { -->
      
      <!--   # ... -->
      
      <!--   # to grob -->
      
      <!--   gp <- ggplotGrob(p) -->
        
        <!--   facet.panels <- grep("^panel", gp[["layout"]][["name"]]) -->
          
          <!--   empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]])) -->
            
            <!--   empty.facet.panels <- facet.panels[empty.facet.panels] -->
              
              <!--   # establish name of empty panels -->
              
              <!--   empty.facet.panels <- gp[["layout"]][empty.facet.panels, ] -->
                
                <!--   names <- empty.facet.panels$name -->
                  
                  <!--   # example of names: -->
                  
                  <!--   #[1] "panel-3-2" "panel-3-3" -->
                  
                  <!--   # now we just need a simple call to reposition the legend -->
                  
                  <!--   lemon::reposition_legend(p, 'center', panel = names) -->
                  
                  <!-- } -->
  
  <!-- fig3 <- shift_legend2(fig3) -->
  
  <!-- ggsave(plot = fig3,  -->
                
                <!--        'Figures/fig3.png',  -->
                
                <!--        dpi = 600,  -->
                
                <!--        width = 11,height = 5) -->
  
  <!-- ``` -->
  ```


# Macroinvertebrate analysis
```{r load-bug-data, warning=FALSE, echo=TRUE}
data <- readxl::read_excel('Data/Samples_Nick_ana_2024.xlsx',
                           sheet = 'Data') %>% 
  dplyr::mutate(Stream = case_when(Stream == 'Arboleda 30' ~ 'Arboleda-30',
                                   Stream == 'Sura 30' ~ 'Sura-30',
                                   Stream == 'Saltito 60' ~ 'Saltito-60',
                                   Stream == 'Piper' ~ 'Piper-30',
                                   Stream == 'Taconazo 30' ~ 'Taconazo-30'),
                functional_group = case_when(`Functional group` == "Predator" ~ 'Predators',
                                             `Functional group` == "Filters" ~ 'Filterers',
                                             `Functional group` == "Collectors-Gatherers" ~ "Collector-Gatherers",
                                             `Functional group` == "Shredder plant tissue" ~ "Shredder plant tissue",
                                             `Functional group` == "Shredder detritus" ~ "Shredder detritus",
                                             `Functional group` == "Scrapers" ~ "Scrapers",
                                             `Functional group` == "Piercers" ~ "Piercers",
                                             `Functional group` == "Miners" ~ "Miners",
                                             `Functional group` == "Shredder detritus" ~ "Shredder detritus",
                                             `Functional group` == "Endoparasite" ~ "Endoparasite"))

data$Stream <- forcats::fct_relevel(data$Stream,
                                    c('Arboleda-30', 'Sura-30', 'Saltito-60',
                                      'Piper-30', 'Taconazo-30'))

# clean the data
tax_data_clean <- data %>% 
  dplyr::mutate(month = as.numeric(gsub(".*?([0-9]+).*", "\\1", Sample)),
                rep = gsub("\\d+", "", Sample)) %>% 
  dplyr::rename(ffg = functional_group) %>% 
  dplyr::filter(QAQC == 0,
                month != 24) 

fig_s1 <- tax_data_clean %>% 
  filter(!is.na(ffg)) %>% 
  group_by(ffg) %>% 
  summarise(total_ffg = sum(Total, na.rm = TRUE)) %>% 
  ggplot(.,
         aes(x = ffg, 
             y = total_ffg))+
  geom_bar(stat = 'identity')+
  geom_label(aes(label = total_ffg),nudge_y = 0.15)+
  scale_y_log10()+
  labs(x = element_blank(),
       y = 'Total Abundance')+
  coord_flip()+
  theme(axis.text.x = element_text(size = 8))

ggsave(plot = fig_s1,
       'Figures/fig_s1.png',
       dpi = 600,
       height = 4, width = 9)

tot_macros <- sum(tax_data_clean$Total)
threshold <- tot_macros*0.01

# main FFGS: at least 1% of all FFGs identified
main_ffgs <- tax_data_clean %>% 
  dplyr::group_by(ffg) %>% 
  dplyr::summarise(n = sum(Total)) %>% 
  dplyr::filter(n > threshold) %>% 
  dplyr::pull(ffg)

```

```{r bug-stats, warning=FALSE, echo=TRUE}

final_dry_mass <- readr::read_csv('Data/final_dry_mass.csv')

final_dry_mass$site <- dplyr::recode_factor(final_dry_mass$site,
                                            Arb = 'Arboleda-30',
                                            Sur30 = 'Sura-30',
                                            Tito60 = 'Saltito-60',
                                            Piper = 'Piper-30',
                                            Tac = 'Taconazo-30')


macro_density <- tax_data_clean %>% 
  dplyr::filter(!is.na(order)) %>%                               # remove NAs from Order
  dplyr::select(site = Stream, Family, month, rep, Total) %>%     # get the necessary columns
  dplyr::group_by(site, month, rep) %>%                          # and do the grouping
  dplyr::summarise(total_family = sum(Total)) %>% 
  dplyr::left_join(final_dry_mass,
                   by = c('site', 'month', 'rep')) %>% 
  dplyr::mutate(abund_per_dm = total_family/dry_mass)

macro_density_sum <- macro_density %>% 
  group_by(site, month) %>% 
  summarise(mean_den = mean(abund_per_dm, na.rm = TRUE),
            se_den = sd(abund_per_dm, na.rm = TRUE)/length(mean_den))

fig3 <- ggplot(macro_density_sum %>%
                 mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))),
               aes(x = month,
                   y = mean_den,
                   color = site,
                   group = site))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin = mean_den - se_den,
                    ymax = mean_den + se_den),
                width = 0)+
  scale_x_continuous(breaks = c(0, 1, 3, 6, 12))+
  labs(y = 'Mean macroinvertebrate density (ind./g DM',
       x = 'Months in stream')+
  scale_color_viridis_d(name = 'Site')+
  theme(axis.title = element_text(size = 12),
        panel.grid = element_blank())+
  theme_classic()
fig3


ggsave(plot = fig3, 
       'Figures/fig3.png', 
       dpi = 600, 
       width = 6, height = 4)


macro_density_ffg <- tax_data_clean %>% 
  dplyr::filter(!is.na(order)) %>%                               # remove NAs from Order
  dplyr::select(site = Stream, ffg, month, rep, Total) %>%     # get the necessary columns
  dplyr::group_by(site, month, ffg, rep) %>%                          # and do the grouping
  dplyr::summarise(total_family = sum(Total))

macro_density_sum_ffg <- macro_density_ffg %>% 
  group_by(site, month, ffg) %>% 
  summarise(mean_abund = mean(total_family, na.rm = TRUE),
            se_abund = sd(total_family, na.rm = TRUE)/length(mean_abund))

fig4 <- ggplot(macro_density_sum_ffg %>% 
                 dplyr::filter(!is.na(ffg),
                               ffg %in% main_ffgs),
               aes(x = month,
                   y = mean_abund,
                   color = site))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin = mean_abund - se_abund,
                    ymax = mean_abund + se_abund),
                width = 0)+
  # geom_boxplot()+
  facet_wrap(ffg ~ .,
             scales = 'free_y')+
  scale_color_viridis_d(name = 'Site')+
  scale_x_continuous(breaks = c(1, 3, 6, 12))+
  scale_y_continuous(breaks = function(y) {
    max_value <- max(y, na.rm = TRUE)
    tick_marks <- seq(0, round(max_value*.75, -1), length.out = 3)
    return(tick_marks)})+
  labs(y = 'Mean FFG Abundance (no.)',
       x = 'Months in stream')+
  theme_classic()+
  theme(axis.title = element_text(size = 12),
        panel.grid = element_blank(),
        strip.background = element_blank())
fig4

# fig4_b <- shift_legend2(fig4_b)

ggsave(plot = fig4, 
       'Figures/fig4.png', 
       dpi = 600, 
       width = 6, height = 4)


macro_density_mod <- macro_density %>% 
  dplyr::mutate(log_abund = log10(abund_per_dm)) %>% 
  dplyr::filter(is.finite(log_abund))

macro_lm <- lm(data = macro_density_mod %>% 
                 dplyr::filter(is.finite(abund_per_dm)),
               log_abund ~ factor(month) * site)

summary(macro_lm)
anova(macro_lm)
hist(macro_lm$residuals)
(agricolae::HSD.test(macro_lm, 'site'))

```

```{r NMDS, warning=FALSE, echo=TRUE}
tax_data_clean_family_log <- tax_data_clean %>% 
  dplyr::filter(!is.na(Family)) %>%                               # remove NAs from Order
  dplyr::select(Stream, Family, month, rep, Total) %>%    # get the necessary columns
  dplyr::group_by(Stream, month, Family) %>%         # and do the grouping
  dplyr::summarise(mean_family = mean(Total, na.rm = TRUE)) %>%   # sum by order in each possible grouping 
  dplyr::mutate(log_total = log10(1 + mean_family)) %>%          # log10 + 1 transform data
  dplyr::select(-mean_family) %>% 
  tidyr::pivot_wider(names_from = Family,                         # pivot data
                     values_from = log_total,
                     values_fill = 0)

tax_data_log_matrix_fam <- tax_data_clean_family_log[,-c(1:2)]
tax_data_log_meta_fam <- tax_data_clean_family_log[,c(1:2)]

nmds_family_log_2d <- vegan::metaMDS(tax_data_log_matrix_fam,
                                     distance = 'bray',
                                     k = 2,
                                     autotransform = FALSE)

stress_fam_log_2d <- round(nmds_family_log_2d$stress, 3)

# evaluate the NMDS via screeplot and Shepard plot
scree <- goeveg::screeplot_NMDS(tax_data_log_matrix_fam, 
                                distance = 'bray', 
                                autotransform = FALSE)

df_scree <- tibble::rownames_to_column(data.frame(scree))

scree_plot <- ggplot(df_scree,
                     aes(x = rowname,
                         y = scree))+
  geom_bar(stat = 'identity')+
  geom_label(aes(label = scree))+
  labs(x = 'Dimensions',
       y = 'Stress')+
  theme_classic()

shepard <- data.frame(vegan::stressplot(nmds_family_log_2d))

shepard_plot <- ggplot(shepard)+
  geom_point(aes(x = x, y = y))+
  geom_line(aes(x = x, y = yf),
            color = 'red')+
  geom_label(x = 0.2, y = 1.5,
             label = 'Non-metric fit R2 = 0.979',
             label.size = 0)+
  geom_label(x = 0.2, y = 1.2,
             label = 'Linear fit R2 = 0.893',
             label.size = 0)+
  labs(x = 'Observed dissimilarity',
       y = 'Ordination distance')+
  theme_classic()+
  theme(panel.grid = element_blank())

fig_s2 <- ggpubr::ggarrange(scree_plot,
                            shepard_plot,
                            labels = 'AUTO',
                            ncol = 2,
                            widths = c(1, 2))

ggsave(plot = fig_s2, 
       'Figures/fig_s2.png', 
       dpi = 600, 
       width = 11, height = 5)


# pick the log transformed 2d NMDS (14.6% stress)

nmds_family_out <- data.frame(x = nmds_family_log_2d$points[,1],
                              y = nmds_family_log_2d$points[,2])

nmds_family_out <- cbind(tax_data_log_meta_fam,
                         nmds_family_out)

fit_fam <- (vegan::envfit(nmds_family_log_2d,
                          tax_data_log_matrix_fam,
                          perm = 9999))

scrs_fam <- data.frame(vegan::scores(fit_fam, 'vectors'))

scrs_fam$pvals <- fit_fam$vectors$pvals

scrs_fam_sig <- subset(scrs_fam, pvals <= 0.05)

scrs_fam_sig$env.variables <- row.names(scrs_fam_sig)


plot_nmds_family_scrs <- ggplot(nmds_family_out,
                                aes(x = x, y = y))+
  geom_point(size = 2)+
  geom_segment(data = scrs_fam_sig,
               aes(x = 0, xend = NMDS1,
                   y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, 'cm')),
               color = 'black')+
  ggrepel::geom_label_repel(data = scrs_fam_sig,
                            aes(NMDS1, NMDS2,
                                label = env.variables),
                            size = 3)+
  labs(x = 'NMDS 1',
       y = 'NMDS 2')+
  lims(x = c(-1,1),
       y = c(-0.75,0.75))+
  theme_classic()+
  theme(panel.grid = element_blank())
plot_nmds_family_scrs


source('Code/veganCovEllipse.R')
streams <- data.frame()
for(i in unique(nmds_family_out$Stream)){
  streams <- rbind(streams,
                   cbind(
                     as.data.frame(
                       with(nmds_family_out[nmds_family_out$Stream == i,],
                            veganCovEllipse(cov.wt(cbind(x, y),
                                                   wt = rep(1/length(x),
                                                            length(x)))$cov,
                                            center = c(mean(x),
                                                       mean(y)
                                            )
                            )
                       )
                     ), 
                     Stream = i)
  )
} # end for loop

plot_nmds_streams <- ggplot(data = nmds_family_out,
                            aes(x = x, y = y))+
  geom_point(aes(color = Stream),
             size = 2)+
  geom_path(data = streams,
            linewidth = 1,
            aes(x = x, y = y, color = Stream))+
  labs(x = "NMDS 1", 
       y = "NMDS 2")+
  lims(x = c(-1,1),
       y = c(-0.75,0.75))+
  scale_color_viridis_d(name = 'Site')+
  geom_label(label = paste('2D stress = ', stress_fam_log_2d),
             x = -0.6,
             y = -0.65,
             label.size = 0)+
  theme_classic()+
  theme(panel.grid = element_blank(),
        legend.background = element_blank(),
        axis.title.y = element_blank())
plot_nmds_streams

fig5 <- ggpubr::ggarrange(plot_nmds_family_scrs,
                          plot_nmds_streams,
                          align = 'h', 
                          widths = c(1,1.25),
                          labels = 'AUTO',
                          label.x = c(0.18,0.13),
                          label.y = 0.98)
fig5

ggsave(plot = fig5, 
       'Figures/fig5.png', 
       dpi = 600, 
       width = 11.5,height = 4)

# Permanova
# The goal of this test is to tell you if there are significant differences in your response variables among your groupings
permanova <- vegan::adonis2(tax_data_log_matrix_fam ~ Stream * month, 
                            data = tax_data_log_meta_fam, 
                            permutations = 9999, 
                            method = 'bray', 
                            by = 'terms')
permanova

# SIMPER
sim <- with(tax_data_log_meta_fam, 
            vegan::simper(tax_data_log_matrix_fam, Stream))

summary(sim)

simper_use <- do.call(rbind, summary(sim)) %>% 
  tibble::rownames_to_column() %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(comparison = strsplit(rowname, "[.]")[[1]][1],
                family = strsplit(rowname, "[.]")[[1]][2]) %>% 
  dplyr::select(-rowname)

simper_sig <- simper_use %>% 
  dplyr::filter(p <= 0.05) %>% 
  dplyr::group_by(comparison, family) %>% 
  dplyr::summarise(n_sig = length(p),
                   mean_avg = mean(average, na.rm = TRUE)) 


table_s2 <- simper_use %>% 
  dplyr::filter(p <= 0.05) %>% 
  dplyr::group_by(comparison, family)

table_s2[,-c(8:9)] <- round(table_s2[,-c(8:9)],
                            digits = 4)
readr::write_csv(table_s2,
                 'Data/table_s2.csv')


# ANOSIM
anosim <- vegan::anosim(tax_data_log_matrix_fam,
                        grouping = tax_data_log_meta_fam$Stream,
                        distance = 'bray',
                        permutations = 9999)
summary(anosim)
anosim_r <- anosim$statistic
anosim_p <- anosim$signif
```

# Session Info

```{r session-info}
pander(sessionInfo())
```
