---
title: "Data for: "
author: "Nicholas S. Marzolf"
date: "Updated March 13, 2023"
output: 
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
RMarkdown file to accompany []. 

To recreate this HTML, load the data files into a folder named 'Data' at the same path as the .rmd file, open the .rmd file, and click 'Knit'.


# Load packages
```{r load-packages, echo = TRUE, message=FALSE, warning = FALSE}
# data manipulation
library(tidyverse)
library(readxl)
library(dplyr)
library(forcats)
library(purrr)

# plotting
library(ggplot2)
library(ggeffects)
library(grid)
library(ggpubr)
library(ggrepel)
library(lemon)
source('C:/Users/Nick Marzolf/Desktop/Research/R code/theme_nick.R')
theme_set(theme_nick())

# statistics
library(car)
library(drc)
library(nlme)
library(nlstools)
library(qpcR)

# spatial
library(sf)
library(raster)
library(spData)
library(tmap)
library(leaflet)
library(spDataLarge)

# community
library(vegan)
library(pander)

```

# Load data

```{r load-data, warning=FALSE, echo=TRUE}
chem <- readxl::read_excel('Data/LT_CWD_datasheet.xlsx',
                           sheet = 'Chemistry') 

cwd <- readxl::read_excel('Data/LT_CWD_datasheet.xlsx',
                          sheet = 'Sheet1')
```

# Stream chemistry for 5 sites
``` {r chemistry, warning=FALSE, echo=TRUE}
chem <- chem %>% 
  dplyr::rename(site = 'Site',
                srp = `SRP (ug/L)`,
                no3_n = `NO3-N (ug/L)`,
                nh4_n = `NH4-N (ug/L)`)

# calculate mean of all measurements
chem_sum <- chem %>% 
  dplyr::group_by(site) %>%
  dplyr::summarise(dplyr::across(.cols = 2:7,
                                 .fns = mean, na.rm = TRUE))

# calculate SD from stream chemistry data
chem_sd <- chem %>% 
  dplyr::group_by(site) %>%
  dplyr::summarise(dplyr::across(.cols = 2:7,
                                 .fns = sd, na.rm = TRUE))

# create object that sorts sites by decreasing mean conductivity
sites <- chem_sum %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean_cond = mean(Cond, na.rm = TRUE)) %>%
  dplyr::arrange(desc(mean_cond)) 

# create a characeter vector of sites in order
sites <- as.character(sites$site)

# create a vector for longer names used in plotting
sites_long <- c(`Arb` = 'Arboleda 30',
                `Sur30` = 'Sura 30',
                `Tito60` = 'Saltito 60',
                `Piper` = 'Piper',
                `Tac` = 'Taconazo 30')


# fill in NA from Arb and Tac with long-term means
# the monthly Arb and Tac data from those sites isn't available yet
chem_sum[1,2] = 201.8
chem_sum[1,3] = 200.0
chem_sum[1,4] = 31.3

chem_sum[4,2] = 3.92
chem_sum[4,3] = 175.4
chem_sum[4,4] = 37.9

# calculate DIN and N:P ratio
chem_sum <- chem_sum %>%
  mutate(din = no3_n + nh4_n,
         n_p = (din/14.0067)/(srp/30.973762))


# re-level the site factor
chem_sum$site <- forcats::fct_relevel(chem_sum$site,
                                      sites)

table1 <- dplyr::arrange(chem_sum, desc(Cond))
table1
```



# Figure 1: Map
```{r map, warning=FALSE, echo=TRUE}
streams_gps <- readr::read_csv('C:/Users/Nick Marzolf/Desktop/NCSU/STREAMS/La Selva GIS data/LTREB Data/GPS sites.csv')


cwd_coords <- streams_gps %>% 
  dplyr::filter(Site %in% sites_long) %>% 
  sf::st_as_sf(., 
               coords = c('Long', 'Lat'),
               crs = st_crs(4326))

cwd_coords$Site <- forcats::fct_relevel(cwd_coords$Site,
                                        c('Arboleda 30', 'Sura 30', 'Saltito 60',
                                          'Piper', 'Taconazo 30'))

# La Selva boundary
lsbs <- sf::st_read(dsn = 'Data/Spatial/laselvaboundary.shp')

# Streams at La Selva shapefile
streams <- sf::st_read(dsn = 'Data/Spatial/streamsclip.shp')


cr <- world %>%
  dplyr::filter(name_long == 'Costa Rica')

# make a basic map of Costa Rica
cr_map <- tmap::tm_shape(cr)+   # create shape based on Costa Rica object
  tmap::tm_polygons()+          # add cr as a polygon
  tmap::tm_shape(lsbs)+         # create shape for the boundary of La Selva
  tmap::tm_dots(size = 1)       # add lsbs as a dot
cr_map


# La Selva boundary and stream layer ----

# Create a map of La Selva boundary and the stream network
map_lsbs <- tmap::tm_shape(lsbs)+            # new shape: La Selva boundary
  tmap::tm_borders()+                        # add as a border/line layer
  tmap::tm_shape(streams)+                   # new shape: stream network
  tmap::tm_lines(col = 'blue')+              # add as a line, colored blue
  tmap::tm_scale_bar(breaks = c(0, 1, 2),    # add a scale bar, with demarkations for 0, 1, and 2 km
                     text.size = 0.75,       # change text size
                     position = c('left',    # put the scale bar in the bottom left
                                  'bottom'))
map_lsbs


fig1 <- tmap::tm_shape(lsbs)+                   # create La Selva boundary layer
  tmap::tm_borders(col = 'black')+
  tmap::tm_shape(streams)+                              # create stream network layer
  tmap::tm_lines(col = 'blue')+
  tmap::tm_shape(cwd_coords)+                             # map the locations of pH sites
  tmap::tm_symbols(size = 1,                            # change the size
                   col = 'Site',                        # colored by site name
                   border.col = 'black',                # with black boundary color
                   palette = "viridis", n = 5)+         # change the color palette
  tmap::tm_scale_bar(breaks = c(0, 1, 2),               # add scale bar
                     text.size = 0.75,
                     position = c('left', 'bottom'))+
  tmap::tm_layout(inner.margins = c(.15,.01, .01, .4),  # change the margins to fit the legend and inset map
                  legend.position = c('right', 'top'),
                  legend.frame = TRUE)+
  tmap::tm_compass(position = c('left', 'top'))+        # add compass north star
  tmap::tm_grid(projection = 4326,
                labels.inside.frame = FALSE,
                lines = FALSE)
fig1

```



# Calculate wood decomposition rates
```{r kCWD, warning=FALSE, echo=TRUE}
# begin cleaning data
cwd <- cwd %>% 
  filter(Flag == 0) %>%
  dplyr::select(site = Site, 
                month = `Collection Month`,
                rep = Rep,
                init_mass = `initial CWD mass (g)`,
                dry_mass = `CWD Pack Dry Mass (g)`,
                init_den = `init wood density (g/cm3)`,
                fin_den = `final wood density (g/cm3)`) %>%
  dplyr::filter(month < 24)

# re-level site factor based on conductivity
cwd$site <- forcats::fct_relevel(cwd$site, 
                                 sites)

# calculate % dry mass remaining
cwd_calc <- cwd %>%
  dplyr::filter(site != 'Sac') %>%
  dplyr::mutate(percent_mass = (dry_mass/init_mass)*100)

# dry mass of each pack to be used in the bug analysis
final_dry_mass <- cwd_calc %>% 
  dplyr::select(site, month, rep, dry_mass) %>% 
  dplyr::filter(month != 0)

# calculate decay rates using ANCOVA
k_cwd_int <- lm(data = cwd_calc,
                log(percent_mass) ~ month * site)

summary(k_cwd_int)
anova(k_cwd_int)
car::Anova(k_cwd_int, type = 'III')

table2 <- cwd_calc %>%
  dplyr::filter(site != 'Sac') %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(
    int = round((coef(lm(log(percent_mass) ~ month))[1]), 3),
    k_yr = (round((coef(lm(log(percent_mass) ~ month))[2])*12, 3))*-1,
    error = summary(lm(log(percent_mass) ~ month))$coefficient[3],
    #df = summary(lm(log(percent_mass) ~ month))$fstatistic,
    r2 = summary(lm(log(percent_mass) ~ month))$r.squared,
    p = anova(lm(log(percent_mass) ~ month))$'Pr(>F'[1])
table2

fig2 <- ggplot(data = cwd_calc,
               aes(y = log(percent_mass),
                   x = month))+
  geom_point(aes(color = site))+
  geom_smooth(aes(color = site),
              method = 'lm',se = FALSE)+
  geom_label(data = table2 %>% 
               dplyr::select(site, k_yr),
             aes(x = 2.5, y = 3.6,
                 label = paste0('k = ',k_yr)))+
  facet_grid(. ~ site, 
             labeller = as_labeller(sites_long))+
  labs(x = 'Months',
       y = 'ln(% DM Remaining)')+
  scale_color_viridis_d()+
  scale_x_continuous(breaks = c(0, 1, 3, 6, 12))+
  theme(legend.position = 'none')
fig2
```



# Merge chemistry with decay rates
```{r decay-chemistry, warning=FALSE, echo=TRUE}
merged <- table2 %>%
  dplyr::select(site, k_yr, error) %>% 
  dplyr::right_join(chem_sum, 'site') %>%
  dplyr::select(-Temp) %>%
  tidyr::pivot_longer(srp:n_p)

names_long <- c(`Cond` = 'Conductivity (µS/cm)',
                `din` = 'DIN (µg/L)',
                `n_p` = 'DIN:SRP',
                `nh4_n` = 'NH4-N (µg/L)',
                `no3_n` = 'NO3-N (µg/L)',
                `pH` = 'pH',
                `srp` = 'SRP (µg/L)')
```

## Run linear and non-linear models of kCWD ~ chemistry
```{r linear-mods, warning=FALSE, echo=TRUE}
vars <- unique(merged$name)

# linear: Y ~ a + bX

# create empty dataframe to populate in the for loop
linear <- data.frame(mod = character(),
                     var = character(),
                     a = numeric(),
                     a_2.5 = numeric(),
                     a_97.5 = numeric(),
                     b = numeric(),
                     b_2.5 = numeric(),
                     b_97.5 = numeric(),
                     AIC = numeric(),
                     p = numeric(),
                     rse = numeric(),
                     df = numeric())


# for loop across each parameter
for(i in 1:length(vars)){
  # which var is being predicted
  var <- vars[i]
  
  # run the lm of k ~ chemistry
  lm <- lm(data = merged %>%
             filter(name == var),
           k_yr ~ value)
  
  # populate the empty data frame
  linear <- linear %>% 
    add_row(
      mod = 'Linear',                    # what kind of model
      var = var,                         # for which variable
      a = coef(lm)[1],                   # intercept
      a_2.5 = confint(lm)[1,1],          # intercept low CI
      a_97.5 = confint(lm)[1,2],         # intercept high CI
      b = coef(lm)[2],                   # slope
      b_2.5 = confint(lm)[2,1],          # slope low CI
      b_97.5 = confint(lm)[2,2],         # slope high CI
      AIC = AIC(lm),                     # model AIC
      p = summary(lm)$coefficients[2,4], # p value
      rse = summary(lm)$sigma,           # RSE
      df = summary(lm)$df[2]             # df
    )
}
```

```{r Mic-Men-mods, warning=FALSE, echo=TRUE}
# empty data frame
micmen <- data.frame(mod = character(),
                     fit = character(),
                     var = character(),
                     a = numeric(),
                     a_2.5 = numeric(),
                     a_97.5 = numeric(),
                     b = numeric(),
                     b_2.5 = numeric(),
                     b_97.5 = numeric(),
                     AIC = numeric(),
                     p = numeric())

# for loop for each variable
for(i in 1:length(vars)) {
  var <- vars[i]
  mm_mod <- try(nls(data = merged %>% 
                      filter(name == var),
                    k_yr ~ SSmicmen(value, a, b))
  )
  
  # fault tolerance, won't work for every parameter
  if(inherits(mm_mod, 'try-error')){
    micmen <- micmen %>% 
      add_row(fit = 'error',
              var = var,
              mod = 'Michaelis-Menten')
    next
  }
  
  # populate the data frame
  micmen <- micmen %>% 
    add_row(
      mod = 'Michaelis-Menten',
      fit = 'success',
      var = var,
      a = coef(mm_mod)[1],
      a_2.5 = confint2(mm_mod, level = 0.95)['a',1],
      a_97.5 = confint2(mm_mod, level = 0.95)['a',2],
      b = coef(mm_mod)[2],
      b_2.5 = confint2(mm_mod, level = 0.95)['b',1],
      b_97.5 = confint2(mm_mod, level = 0.95)['b',2],
      AIC = AIC(mm_mod),
      p = summary(mm_mod)$coefficients[8]
    )
}
```

```{r logarithmic, warning=FALSE, echo=TRUE}
## Logarithmic: Y ~ a + b*log(X)
logarithmic <- data.frame(mod = character(),
                          var = character(),
                          a = numeric(),
                          a_2.5 = numeric(),
                          a_97.5 = numeric(),
                          b = numeric(),
                          b_2.5 = numeric(),
                          b_97.5 = numeric(),
                          AIC = numeric(),
                          R2 = numeric(),
                          p = numeric(),
                          rse = numeric(),
                          df = numeric()
)

for(i in 1:length(vars)) {
  var <- vars[i]
  
  loga <- lm(data = merged %>% 
               filter(name == var),
             k_yr ~ log10(value)) 
  
  logarithmic <- logarithmic %>% 
    add_row(
      mod = 'Logarithmic',
      var = var,
      a = coef(loga)[1],
      a_2.5 = confint(loga)[1],
      a_97.5 = confint(loga)[3],
      b = coef(loga)[2],
      b_2.5 = confint(loga)[2],
      b_97.5 = confint(loga)[4],
      AIC = AIC(loga),
      R2 = summary(loga)$r.squared,
      p = summary(loga)$coefficients[8],
      rse = summary(loga)$sigma,
      df = summary(loga)$df[2]
    )
}
```

```{r logistic, warning=FALSE, echo=TRUE}
## Logistic: Y ~ 1/1+exp(X)
logistic <- data.frame(mod = character(),
                       var = character(),
                       b = numeric(),
                       b_2.5 = numeric(),
                       b_97.5 = numeric(),
                       d = numeric(),
                       d_2.5 = numeric(),
                       d_97.5 = numeric(),
                       e = numeric(),
                       e_2.5 = numeric(),
                       e_97.5 = numeric(),
                       AIC = numeric(),
                       rse = numeric(),
                       df = numeric())

for(i in 1:length(vars)){
  var <- vars[i]
  
  logi <- drm(k_yr ~ value, 
              data = merged %>% 
                filter(name == var),
              fct = L.3())
  
  logistic <- logistic %>% 
    add_row(
      mod = 'Logistic',
      var = var,
      b = coef(logi)[1],
      b_2.5 = confint2(logi)['b:(Intercept)',1],
      b_97.5 = confint2(logi)['b:(Intercept)',2],
      d = coef(logi)[2],
      d_2.5 = confint2(logi)['d:(Intercept)',1],
      d_97.5 = confint2(logi)['d:(Intercept)',2],
      e = coef(logi)[3],
      e_2.5 = confint2(logi)['e:(Intercept)',1],
      e_97.5 = confint2(logi)['b:(Intercept)',2],
      AIC = AIC(logi),
      rse = summary(logi)$rseMat[1],
      df = summary(logi)$rseMat[2])
}

```

## Combine and evaluate model fits
```{r model-comps, warning=FALSE, echo=TRUE}
# combine model outputs

aic_all <- rbind(logistic %>% 
                   dplyr::select(mod, var, AIC) %>% 
                   mutate(p = NA),
                 logarithmic %>% 
                   dplyr::select(mod, var, AIC, p),
                 linear %>% 
                   dplyr::select(mod, var, AIC, p),
                 micmen %>% 
                   dplyr::select(mod, var, AIC, p))

best_fits <- aic_all %>% 
  arrange(var,
          AIC) %>% 
  group_by(var) %>% 
  dplyr::slice_head(n=1)


aic_wts <- data.frame()
for(i in 1:length(vars)) {
  use <- vars[i]
  
  df <- aic_all %>% 
    filter(!is.na(AIC),
           var %in% use) %>% 
    mutate(delAIC = akaike.weights(AIC)$deltaAIC,
           weights = akaike.weights(AIC)$weights)
  
  aic_wts <- rbind(aic_wts, df)
}


table3 <- aic_wts %>% 
  dplyr::select(var, mod, weights) %>% 
  pivot_wider(names_from = var, 
              values_from = weights) %>% 
  knitr::kable(format = 'rst')
table3

```


```{r pred-k, warning=FALSE, echo=TRUE}
# plot decay rates as a function of chemistry
# add lines for SRP (linear), NH4 (M-M), N:P(M-M), and cond (linear)

merged_preds <- rbind(merged %>% 
                        filter(name == 'srp') %>% 
                        mutate(pred_k = -0.12456 + (-0.002889*value)),
                      merged %>% 
                        filter(name == 'Cond') %>% 
                        mutate(pred_k = -0.075 + (-0.0022*value)) ,
                      merged %>% 
                        filter(name == 'nh4_n') %>% 
                        mutate(pred_k = (-0.0768*value)/(-28.13+value)) ,
                      merged %>% 
                        filter(name == 'n_p') %>% 
                        mutate(pred_k = (-0.179*value)/(-1.944+value)),
                      merged %>% 
                        filter(name %in% c('no3_n', 'din', 'pH')) %>% 
                        mutate(pred_k = NA))

merged_preds <- data.frame()
for(i in 1:length(best_fits$var)) {
  fit <- best_fits[i,]
  
  type <- pull(fit[1])
  var <- pull(fit[2])
  
  if(type == 'Logistic'){
    logis <- logistic %>% 
      filter(var == !!var)
    
    df <- merged %>% 
      filter(name == var) %>% 
      mutate(pred_k = logis$d/(1+exp(logis$b*(value - logis$e))),
             best_fit = type)
  
  } # end logistic if statement
  
  if(type == 'Linear') {
    line <- linear %>% 
      filter(var == !!var)
    
    if(line$p <= 0.05){
    df <- merged %>% 
      filter(name == var) %>% 
      mutate(pred_k = line$a + (value*line$b),
             best_fit = type)
    } else {
      df <- merged %>% 
        filter(name == var) %>% 
        mutate(pred_k = NA,
               best_fit = NA)
    }
  } # end linear if statement
  
  if(type == 'Michaelis-Menten'){
    mm <- micmen %>% 
      filter(var == !!var)
    
    if(mm$p <= 0.05){
    df <- merged %>% 
      filter(name == var) %>% 
      mutate(pred_k = (mm$a*value)/(value + mm$b),
             best_fit = type)
    } else {
      df <- merged %>% 
        filter(name == var) %>% 
        mutate(pred_k = NA,
               best_fit = NA)
    }
  } # end M-M for loop
  
  if(type == 'Logarithmic'){
    logar <- logarithmic %>% 
      filter(var == !!var)
    
    if(logar$p <= 0.05) {
      df <- merged %>% 
        filter(name == var) %>% 
        mutate(pred_k = logar$a + (logar$b*value),
               best_fit = type)
    } else {
      df <- merged %>% 
        filter(name == var) %>% 
        mutate(pred_k = NA,
               best_fit = NA)
    }
  } # end logarithmic if statement
  
  merged_preds <- rbind(merged_preds, df)
} # end for loop


fig3 <- ggplot(merged_preds)+
  geom_point(aes(x = value,
                 y = k_yr,
                 color = site),
             size = 4)+
  geom_line(aes(x = value, 
                y = pred_k),
            size = 1,
            linetype = 'dashed')+
  ylab(expression(paste(k[CWD], ' (', y^-1,')')))+
  facet_wrap(. ~ name,
             nrow = 2,
             scales = 'free',
             labeller = as_labeller(names_long),
             drop = FALSE)+
  scale_color_viridis_d(name = 'Site',
                        labels = as_labeller(sites_long))+
  theme(axis.title.x = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))
fig3

shift_legend2 <- function(p) {
  # ...
  # to grob
  gp <- ggplotGrob(p)
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]

  # establish name of empty panels
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  names <- empty.facet.panels$name
  # example of names:
  #[1] "panel-3-2" "panel-3-3"

# now we just need a simple call to reposition the legend
  lemon::reposition_legend(p, 'center', panel=names)
}


shift_legend2(fig3)
```


# Macroinvertebrate analysis

```{r load-bug-data, warning=FALSE, echo=TRUE}
data <- readxl::read_excel('Data/Samples_Nick.xlsx')
taxonomy <- readxl::read_excel('Data/Samples_Nick.xlsx',
                               sheet = 'taxonomy')

data$Stream <- forcats::fct_relevel(data$Stream,
                           c('Arboleda 30', 'Sura 30', 'Saltito 60',
                             'Piper', 'Taconazo 30'))

# build taxonomy for each ID
tax_data <- list()
for(i in 1:length(unique(data$Taxa))){
  rank <- unique(data$Taxa)[i]
  
  tax_info <- taxonomy %>% 
    dplyr::filter(Taxa %in% rank)
  
  tax_data[[i]] <- data %>% 
    data.frame() %>% 
    dplyr::filter(Taxa == rank) %>% 
    dplyr::mutate(rank = tax_info$Rank,
           family = tax_info$Family,
           order = tax_info$Order,
           class = tax_info$Class,
           phylum = tax_info$Phylum,
           month = as.numeric(gsub(".*?([0-9]+).*", "\\1", Sample)),
           rep = gsub("\\d+", "", Sample))
}

# create df
tax_data <- purrr::reduce(tax_data, rbind)

# clean the data
tax_data_clean <- tax_data %>% 
  dplyr::filter(QAQC == 0,
         month != 24) 
```

```{r bug-stats, warning=FALSE, echo=TRUE}

final_dry_mass <- readr::read_csv('Data/final_dry_mass.csv')

final_dry_mass$site <- recode_factor(final_dry_mass$site,
                                     Arb = 'Arboleda 30',
                                     Sur30 = 'Sura 30',
                                     Tito60 = 'Saltito 60',
                                     Piper = 'Piper',
                                     Tac = 'Taconazo 30')


macro_density <- tax_data_clean %>% 
  dplyr::filter(!is.na(order)) %>%                               # remove NAs from Order
  dplyr::select(site = Stream, order, month, rep, Total) %>%     # get the necessary columns
  dplyr::group_by(site, month, rep) %>%                          # and do the grouping
  dplyr::summarise(total_order = sum(Total)) %>% 
  dplyr::left_join(final_dry_mass,
            by = c('site', 'month', 'rep')) %>% 
  dplyr::mutate(abund_per_dm = total_order/dry_mass)

macro_density_sum <- macro_density %>% 
  group_by(site, month) %>% 
  summarise(mean_den = mean(abund_per_dm, na.rm = TRUE),
            se_den = sd(abund_per_dm, na.rm = TRUE)/length(mean_den))

figure4 <- ggplot(macro_density_sum,
       aes(x = month,
           y = mean_den,
           color = site,
           group = site))+
  geom_point()+
  geom_line()+
  geom_errorbar(aes(ymin = mean_den - se_den,
                    ymax = mean_den + se_den),
                width = 0)+
  scale_x_continuous(breaks = c(0, 1, 3, 6, 12))+
  labs(y = expression(paste('Mean macroinvertebrate density (# g   ',DM^-1,')')),
       x = 'Month')+
  scale_color_viridis_d(name = 'Site')+
  theme(axis.title = element_text(size = 12))
figure4


macro_lm <- lm(data = macro_density,
               log10(abund_per_dm) ~ factor(month) * site)

summary(macro_lm)
anova(macro_lm)
hist(macro_lm$residuals)
(agricolae::HSD.test(macro_lm, 'site'))

```

```{r NMDS, warning=FALSE, echo=TRUE}

```



# Session Info
```{r session-info}
pander(sessionInfo())
```
































